
```{r}
rm(list=ls())
```

#Install the rentrez package to query the downloaded sequences
```{r}
# install.packages("rentrez")
# install.packages("tidyverse")
# install.packages("cli")
install.packages("svglite")
# library(rentrez)
library(tidyverse)
library(dplyr)
library(svglite)
```

#Create heatmap.
First gather all the relevant files together

```{r}
library(readr)

folder_append <- function(pattern_match, starting_directory){

  # List all .tsv files in the subdirectory
  files <- list.files(path = starting_directory, pattern = pattern_match, full.names = TRUE, recursive = TRUE)

  # Load and combine all .tsv files into one dataframe
  df <- do.call("rbind", lapply(files, function(x) {
    data <- read_tsv(x, show_col_types = FALSE)
    data$file_name <- basename(x)  # Add the file name as a new column
    return(data)
  }))
  
  return(df)
}

#Also get the GDBTK File
# gtdb_data <- read_tsv("C:\\Users\\Justin\\Desktop\\Bioinf_PBI\\gtdbtk_summary.tsv")

#Add the list of controls
control_list <- read.csv("control_list.csv")

```


```{r}
#Test on one folder
#test_1 <- folder_append("product_.*\\.tsv","C:\\Users\\Justin\\Desktop\\Bioinf_PBI\\DRAM_controls\\gather_dram_product_tsv\\")

#Function to import folder
dram_append <- folder_append("product_.*\\.tsv",
                        "C:\\Users\\Justin\\Desktop\\Bioinf_PBI\\DRAM gathered - 23 July")


```


```{r}
#Tidy the dram file
dram_clean <- dram_append %>%
  #Add a column to extract between the first two underscores
  mutate(sample = sub("product_","",file_name)) %>%
  #If the file sample starts with GCA
  mutate(sample = ifelse(substr(sample,1,3)=="GCA",
                            #Then extract the GCA bit to the next underscore
                            gsub(".*(GCA_[^_]*).*", "\\1", sample),
                            #Otherwise just extract to the first underscore
                            gsub("_.*$", "\\1", sample))) %>%
  #Exract the Binner from this
  rowwise()%>%
  mutate(Binner=gsub(paste0("product_",sample,"_"),"",file_name),
         Binner=gsub(paste0("_",genome,".tsv"),"",Binner))%>%
  ungroup() %>%
  #Then just clear anything which was downloaded from NCBI, rather than binned
  mutate(Binner= ifelse(substr(sample,1,3)=="GCA","",Binner))%>%
  select(sample,genome,Binner,everything())

#Exclude columns in gtdb merge
gtdb_merge <- gtdb_data %>%
  select(user_genome,classification,Sample,Binner)

#Merge with GTDB
dram_gtdb <- dram_clean %>%
  left_join(gtdb_merge,by=c("sample"="Sample","Binner"="Binner","genome"="user_genome")) %>%
  select(classification,everything()) %>%
  left_join(control_list, by=c("sample"="sample")) %>%
  mutate(classification=ifelse(is.na(classification),Organism,classification))

write.csv(dram_gtdb,"dram_controls_csp1-3.csv", row.names = FALSE,na="")
```

#Clean so that can create heatmap
#Careful if there's two specimens with metabat sspec and ssens as these will be grouped in so might need to use alternate code below:
(Maybe make that dynamic if going to be re-running)
```{r}
#
heatmap_tidy <- dram_gtdb %>%
  #Get rid of the DSM bits after the controls
  mutate(Organism = gsub("7 = DSM 20455|DSM \\d*$","",classification)) %>%
  #Replace all SRA GTDB identified specimens with CSP1-3 then the sample and binner
  rowwise() %>%
  mutate(Organism = gsub("^d__.*CSP.*",paste0("CSP1-3 ",sample," ",Binner," ",genome),Organism)) %>%
  ungroup()%>%
  mutate(Organism = gsub("_bins","",Organism))%>% 
  #Have CSP1-3 starting with the GCA data so it arranges with the SRA CSP1-3
  mutate(Organism = gsub("Armatimonadetes bacterium CSP1-3","CSP1-3 Armatimonadetes bacterium",Organism))%>% 
  mutate(Organism = gsub("Bacterium HR32","CSP1-3 Bacterium HR32",Organism)) %>%
  #If it's a control then add that to the organism name
  rowwise()%>%
  mutate(Organism = ifelse(substr(sample,1,3)=="GCA",paste(Organism,sample),Organism))%>%
  ungroup() %>%
  #Select the organism and relevent columns
  select(Organism,5:17)

```

#just try from here...
```{r}
merge_data <- read.csv("C:\\Users\\judavis\\Desktop\\Bioinf\\merged_data_1448_07_10_2023_with_drep_org.csv", check.names = FALSE, stringsAsFactors = FALSE)

```


```{r}

# Extract the new column names from the first row
new_col_names <- as.character(unlist(merge_data[1, ]))

# Remove the first row from the original data frame
heatmap_data <- merge_data[-1, ]

# Assign the new column names
colnames(heatmap_data) <- new_col_names

# Convert all columns to characters for consistency
heatmap_data[] <- lapply(heatmap_data, as.character)

# Now, since you have exact column names, use backticks for non-standard names in dplyr operations
library(dplyr)

heatmap_tidy <- heatmap_data %>%
  filter(Drep == "Yes") %>%
  select(Organism, 215: 226) %>%
  mutate(across(-Organism, as.numeric, .names = "{.col}"))

```




```{r}
# Load the required libraries
library(tidyverse)
library(reshape2) # for melt function
library(viridis) # for viridis color palette
library(forcats) #To fix the stupid catagorical variables not getting alphabetised on the heatmap

# Transform it into long format
heatmap_long <- melt(heatmap_tidy, id.vars = "Organism") %>%
  arrange %>%
  mutate(value = as.numeric(value))

# Convert value to percentage and round
heatmap_long$value <- round(heatmap_long$value * 100) 

#Arrange it
heatmap_long <- heatmap_long %>% 
  #Hydroxypropionate is messing up the formatting of the chart
  mutate(variable=gsub("Hydroxypropionate","Hydroxy-propionate",variable))%>%
  #Put a space so hopefully the wordwrap works beter
  mutate(variable=gsub("-","- ",variable))%>%
  arrange(Organism)

#Need to have labels as factors or it won't order alphabetically...
# Convert 'Organism' to a factor and order its levels alphabetically
heatmap_long$Organism <- factor(heatmap_long$Organism, levels = sort(unique(heatmap_long$Organism)))
```


Heatmap #1 labels at the bottom
```{r}
# Create the color palette
color_palette <- colorRampPalette(c("white", "purple"))(100)

# Create the heatmap
heatmap_image <- ggplot(heatmap_long, aes(x = variable, y = Organism, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = color_palette) +
  theme_minimal() +
  labs(x = "Percentage of Modules identified by DRAM for each Pathway", y = "Organism") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_text(aes(label = paste0(value, "%")), color = "black", size = 4)

# Save the heatmap to a image files
ggsave(filename = "heatmap.png", plot = heatmap_image, width = 15, height = 15, dpi = 300)
ggsave(filename = "heatmap.svg", plot = heatmap_image, width = 15, height = 15)
```

```{r}
heatmap_image <- ggplot(heatmap_long, aes(x = variable, y = Organism, fill = value))
print(heatmap_image)
```


Heatmap #1 labels at the bottom
```{r}
# Load the required library
library(stringr)

# Create the heatmap
heatmap_image <- ggplot(heatmap_long, aes(x = variable, y = Organism, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = color_palette) +
  scale_x_discrete(position = "top") +
  theme_minimal() +
  labs(x = "Percentage of Modules identified by DRAM for each Pathway", y = "Organism") +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.5)) +
  geom_text(aes(label = paste0(value, "%")), color = "black", size = 4)

# Modify x-axis labels to be wrapped
heatmap_image <- heatmap_image + 
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))

ggsave(filename = "heatmap_2.png", plot = heatmap_image, width = 15, height = 15, dpi = 300)
ggsave(filename = "heatmap_2.svg", plot = heatmap_image, width = 15, height = 15)

```
#Third time's a charm
```{r}
# Load the required library


# Create the heatmap
heatmap_image <- ggplot(heatmap_long, aes(x = variable, y = Organism, fill = value)) +
  geom_tile() +
  scale_fill_gradientn(colours = color_palette) +
  scale_x_discrete(position = "top") +
  theme_minimal() +
  theme(axis.title.x = element_text(vjust = 1),  # move x-axis title to the top
        axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.5)) +
  labs(x = "Percentage of Modules identified by DRAM for each Pathway", y = "Genome") +
  geom_text(aes(label = paste0(value, "%")), color = "black", size = 4)

# Modify x-axis labels to be wrapped
heatmap_image <- heatmap_image + 
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5)) +
  scale_x_discrete(labels = function(x) str_wrap(x, width = 10))

# Print the heatmap
print(heatmap_image)

ggsave(filename = "heatmap_3.png", plot = heatmap_image, width = 15, height = 15, dpi = 300)
ggsave(filename = "heatmap_3.svg", plot = heatmap_image, width = 15, height = 15)
```


```{r}
library(ggplot2)
library(reshape2)
library(data.table)

dt <- read.table("http://archive.ics.uci.edu/ml/machine-learning-databases/wine/wine.data", sep = ",")
cor.mat <- cor(dt[1:10], method = "spearman")
cor.mat.melt <- melt(cor.mat)
colnames(cor.mat.melt) <- c("x1","x2","Corelation")
cor.mat.melt$x1 <- factor(cor.mat.melt$x1, levels = c("V10", "V9", "V8", "V7", "V6", 
                                                      "V1", "V2", "V3", "V4", "V5"))
cor.mat.melt$x2 <- factor(cor.mat.melt$x2, levels = c("V10", "V9", "V8", "V7", "V6", 
                                                      "V1", "V2", "V3", "V4", "V5"))



ggplot(data = cor.mat.melt,
       aes(x = x1, y = x2)) +
geom_tile(aes(fill = Corelation)) +
scale_fill_gradientn(colours = rainbow(3)) +
geom_text(aes(x = x1, y = x2, label = round(Corelation, 2))) +
labs(x = "", y = "")
```


```{r}
print(heatmap_image)
```



```{r}
names(dram_gtdb)
```


#Ye Olde Code
Might need to use this later so don't delete....
```{r}
# library(readr)
# 
# folder_append <- function(pattern_match, starting_directory){
# 
# # List all .tsv files in the subdirectory
# files <- list.files(path = starting_directory, pattern = pattern_match, full.names = TRUE,recursive = TRUE)
# 
# # Load and combine all .tsv files into one dataframe
# df <- do.call("rbind", lapply(files, function(x) read_tsv(x,show_col_types=FALSE)))
#   
# }
```


